Commentary and in-play models treat the death overs (16--20) of Twenty20 (T20)
cricket as a place where ``momentum'' is real: a batter who has just found the
boundary is hot, and a run of dot balls builds pressure that spills into the next
delivery. We show that the most-discussed version of this --- the death-over
``cold hand'' --- is almost entirely a statistical mirage, and we measure its
size. Importing the Miller--Sanjurjo finite-sequence selection correction to
per-ball cricket --- to our knowledge the first time the bias correction itself
has been applied at the delivery level --- we put the apparent-momentum illusion
at $+0.121$: the naive momentum estimate ($D=-0.114$) is almost exactly cancelled
by the finite-sequence bias ($-0.121$), so the corrected, state-controlled
estimate is $+0.007$ and not significant against a state-stratified null
($p=0.40$). Death-over stays are short, the regime where the bias is largest, so
this is precisely where the uncorrected estimator misleads most. We frame this as
a \emph{measurement of how large the mirage is} --- the quantity we are powered to
estimate --- rather than as a test of whether momentum can exist at all.

On 223{,}678 death-over deliveries from 4{,}402 men's IPL and T20I matches
(Cricsheet, 2005--2026) we build calibrated per-ball models of expected runs and
wicket probability, and two further results follow. First, a historical-rate
encoding of batter/bowler identity does not improve out-of-sample skill over a
game-state-only model --- it worsens held-out expected-runs RMSE in every one of
six rolling seasons --- even though SHAP attributes $51$--$64\%$ of explanatory
mass to batter identity; a random-player placebo collapses that mass to $7$--$8\%$
with skill unchanged, a cautionary dissociation between in-sample attribution and
out-of-sample skill. Second, residual dot-ball pressure adds skill to expected
runs (RMSE $+0.0065$, 95\% CI $[0.0048, 0.0083]$; longer dot runs lower expected
runs) but nothing to wicket probability, so the Markov property is violated for
scoring intensity but not dismissal risk --- an associational effect, plausibly
confounded with bowler quality. Per-ball death-over outcomes sit close to the base
rate, so absolute skill gains are small; our claims rest on the bias correction
and the qualitative dissociation, not on the margin of any single model.
